Predicting Remaining Useful Life of NASA Aircraft Engines

### 2025-03-08 index.qmd
minor update

Cox proportional hazards (CPH) models are statistical regression models
specializing in modeling time-to-event predictions with survival data.
Survival data is data with a value for time and an event
[@Abeysekera2009]. For example, in a study investigating the survival
likelihood of mechanical equipment, the time value would begin when the
equipment begins functioning and would conclude in the event the
equipment stops functioning [@Smith2003]. CPH models can use single or
multiple covariates as a means of creating predictions until failure and
are capable of dealing with survival data that contains censored data as
well [@Seung2023]. Censored data is when the information about an
individual in a study is only known for a certain period of time, such
as the length of the study [@klein2005, 1]. The models also have the
advantage of being semi-parametric which allows more flexibility as
compared to other models [@Ming-Chiang2014].

@@

The Kaplan-Meier curve on the X-axis represents the length of survival
time on the interval from 0 to the largest observation time while the
Y-axis represents the cumulative probability of surviving at a given
time. The curve is based on step-wise estimates thus making the curve
non-continuous [@rich2010]. Generating a Kaplan-Meier curve requires the
R packages `survival` and `survminer` and uses the survival object and
survival formula for the data. The below curve was generated using the
`survival` package and dataset "lung" and demonstrates the general
survival probability.

@@

To generate the CPH model, a survival object must first be created.
Using the R package `survival`, the survival formula can be generated by
running `Surv(time, event) ~ x`. That formula is then applied to the R
function `coxph(formula, data=)` which will generated the CPH model
[@R-base].

>>>>

Cox proportional hazards (CPH) models are statistical regression models
specializing in modeling time-to-event predictions with survival data
introduced by D. R. Cox in 1972 [@nahhas2025]. Survival data is data
with a value for time and an event [@Abeysekera2009]. For example, in a
study investigating the survival likelihood of mechanical equipment, the
time value would begin when the equipment begins functioning and would
conclude in the event the equipment stops functioning [@Smith2003]. CPH
models can use single or multiple covariates as a means of creating
predictions until failure and are capable of dealing with survival data
that contains censored data as well [@Seung2023]. Censored data is when
the information about an individual in a study is only known for a
certain period of time, such as the length of the study [@klein2005].
The models also have the advantage of being semi-parametric which allows
more flexibility as compared to other models [@Ming-Chiang2014].

@@

The Kaplan-Meier curve on the X-axis represents the length of survival
time on the interval from 0 to the largest observation time while the
Y-axis represents the cumulative probability of surviving at a given
time. The curve is based on step-wise estimates thus making the curve
non-continuous [@rich2010]. Generating a Kaplan-Meier curve requires the
R packages `survival` and `survminer` and uses the survival object and
survival formula for the data. The below curve was generated using the
`survival` package and dataset "lung" and demonstrates the general
survival probability using the `ggsurvplot` function [@R-base].

@@

To generate the CPH model, a survival object must first be created.
Using the R package `survival`, the survival formula can be generated by
running `Surv(time, event) ~ x`. That formula is then applied to the R
function `coxph(formula, data=)` which will generate the CPH model
[@R-base]. The output of the CPH model includes the coefficients
(\$\\beta\$ and the confidence intervals for each covariate, and the
concordance value, likelihood ratio test, Wald test, and the score
(logrank) test of the overall model [@nahhas2025].
